Using Distributed Information Stores for Patient-Clinician Health Data Sharing

### 2013-01-16 patient-clinician-workshop/pccworkshop.tex
sleep

In this position paper, we look at the potential and problems of
augmenting the next generation of EHRs with \emph{personal activity
  diaries}, portraits of a person's dynamic condition captured at high
temporal fidelity.  Such activity diaries are intended to supplement,
rather than supplant, the existing diagnostic reports performed by
professionals, first and foremost, to help clinicians derive a more
complete picture of the person throughout a period of care, such as an
intervention.  We believe that such activity diaries could potentially
also be beneficial to individuals as well, making people more aware of
the evolution of their health.

@@

product & measurement indices & price range (approx. USD) \\
\hline
FitBit & general activity (walking/jogging) & \$50 \\
Nike FuelBand & general activity (walking/jogging) & \$199 \\ 
Writhings WiFi Scale & weight & \$200 \\
Writhings BPM & blood pressure & \$200 \\ 
Zeo Pro & sleep duration \& quality & \$200 \\
WakeMate & sleep duration & \$50  \\
Zensorium Tinkne & heart rate variability, blood oxygenation, respiratory rate & \$200 \\
BodyMedia CORE 2 & galvanic skin repsonse, body temperature, sleep, respiratory rate, heart rate & (TBD) \\
Hapifork & eating time, rate, frequency & (TBD) \\ 
\end{tabular}
\caption{\emph{Consumer-grade activity sensors}: Activity and physiological measurememnt sensors with automatic data uplink to cloud services }
\label{sensors}
\end{table}
}
\vspace{6cm}

@@

The recent proliferation in personal wearable activity sensors and
network and storage- enabled digital health measurement devices has
been driven by both the falling costs of embeddable electronics, and a
surging public interest in ``quantified-self''-style self-monitoring
for self-improvement.  Many of these sensors are lightweight, can be
worn discreetly, measure several physiological statistics
simultaneously, and most require little maintenance. Table
\ref{sensors} lists a variety of such sensors, currently available
readily from online retailers.

While few of these devices had been independently evaluated for
accuracy for clinical use, as they have matured, several of these
manufacturers begun rigorous evaluaton of their products for achieving
certification. For example, BodyMedia advertise themselves as
providing the ``FDA-approved, clincally validated, most accurate
wireless activity measurement''.  A separate study of activity-based
sleep monitoring tools revealed that ``low cost actigraphy-based
approaches'' correlated well with baseline (in-lab EEG-based
measurements) for sleep duration, although less with sleep quality
metrics.

>>>>

In this position paper, we look at the potential and problems of
augmenting the next generation of EHRs with personal \emph{activity
  diaries}, portraits of a person's dynamic condition captured at high
temporal fidelity.  Such activity diaries are intended to supplement,
rather than supplant, the existing diagnostic reports performed by
professionals, first and foremost, to help clinicians derive a more
complete picture of the person throughout a period of care, such as an
intervention.  We believe that such activity diaries could potentially
also be beneficial to individuals as well, making people more aware of
the evolution of their health.

@@

product & measurement indices & price range (approx. USD) \\
\hline
FitBit & general activity (walking/jogging) & \$50 \\
Nike FuelBand & general activity (walking/jogging) & \$199 \\ 
Writhings WiFi Scale & weight & \$200 \\
Writhings BPM & blood pressure & \$200 \\ 
Zeo Pro & sleep duration \& quality & \$200 \\
WakeMate & sleep duration & \$50  \\
Zensorium Tinkne & heart rate variability, blood oxygenation, respiratory rate & \$200 \\
BodyMedia CORE 2 & galvanic skin repsonse, body temperature, sleep, respiratory rate, heart rate & (TBD) \\
Hapifork & eating time, rate, frequency & (TBD) \\ 
\end{tabular}
\caption{\emph{Consumer-grade activity sensors}: Activity and physiological measurement sensors with automatic data uplink to cloud services }
\label{sensors}
\end{table}
}
\vspace{6cm}

@@

The recent proliferation in personal wearable activity sensors and
network and storage- enabled digital health measurement devices has
been driven by both the falling costs of embeddable electronics, and a
surging public interest in ``quantified-self''-style self-monitoring
for self-improvement.  Many of these sensors are lightweight, can be
worn discreetly, measure several physiological statistics
simultaneously, and most require little maintenance. Table
\ref{sensors} lists a sampling of the sensors currently available
directly from retailers.

While few of these devices had initially been evaluated for accuracy
or certified for clinical use, increasingly device manufacturers have
begun to get independent certification. For example, BodyMedia
advertise themselves as providing the ``FDA-approved, clincally
validated, most accurate wireless activity measurement''.  A separate
study of activity-based sleep monitoring tools revealed that ``low
cost actigraphy-based approaches'' correlated well with baseline
(in-lab EEG-based measurements) for sleep duration, although less with
sleep quality metrics \cite{reid1999correlation}.
